---
title: "An Introduction to the GSDA Package"
author: "xueyuan cao, PhD, Stan Pounds, PhD"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An Introduction to the GSDA Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r loadLibs, include = FALSE}
library(MASS)
library(GSDA)
library(ggplot2)
library(knitr)
data(target.aml.clin)
data(target.aml.expr)
data(kegg.ml.gsets)
opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  digits = 3,
  tidy = FALSE,
  background = "#FFFF00",
  fig.align = 'center',
  warning = FALSE,
  message = FALSE
  )
options(width = 55, digits = 3)
theme_set(theme_bw())
getInfo <- function(what = "Suggests") {
  text <- packageDescription("GSDA")[what][[1]]
  text <- gsub("\n", ", ", text, fixed = TRUE)
  text <- gsub(">=", "$\\\\ge$", text, fixed = TRUE)
  eachPkg <- strsplit(text, ", ", fixed = TRUE)[[1]]
  eachPkg <- gsub(",", "", eachPkg, fixed = TRUE)
  #out <- paste("\\\**", eachPkg[order(tolower(eachPkg))], "}", sep = "")
  #paste(out, collapse = ", ")
  length(eachPkg)
}
```


The **GSDA** package (short for Gene-set Distance Analysis) contains functions to to carry out the gene-set distance analysis (GSDA) by generalizing distance correlations to evaluate the association of a gene set with categorical, quantitative, and censored event-time variables. **GSDA** loads packages as needed and assumes that they are installed. 



An example to run GSDA on target aml RNAseq with Chloroma status (categorical) with myeloid gene sets defined by kegg.
```{r Chloroma}
chl.res=gsda(sqrt(target.aml.expr),
             target.aml.clin,
             kegg.ml.gsets,
             "Chloroma","oe","ct")
chl.res
```

An example to run GSDA on target aml RNAseq with White blood counts (quantitative) with myeloid gene sets defined by kegg.
```{r logWBC}
wbc.res=gsda(sqrt(target.aml.expr),
             target.aml.clin,
             kegg.ml.gsets,
             "logWBC","oe","oe")
wbc.res

```

An example to run GSDA on target aml RNAseq with event free survival (efs, censored event-time) with myeloid gene sets defined by kegg.
```{r efs}
efs.res=gsda(sqrt(target.aml.expr),
             target.aml.clin,
             kegg.ml.gsets,
             c("efs.time","efs.evnt"),"oe","st")
efs.res

```

Install **GSDA** using

```{r install, eval = FALSE}
devtools::install_github('xueyuancao/GSDA', dependencies = c("Depends", "Suggests"))
```